\section{Related Work}

% We discuss related work from three perspectives, namely implicit shape representations (Section~\ref{Subsec:Implicit:Shape:Rep}),  hybrid 3D representations (Section~\ref{Subsec:Hybrid:3D:Rep}), and coordinate field optimization (Section~\ref{Subsec:Coordinate:Field:Opt}).

%\vspace{-0.05in}
\noindent\textbf{Implicit Shape Representations.}
Implicit shape representations are state-of-the-art in encoding shape geometric details~\cite{DBLP:conf/cvpr/ParkFSNL19, DBLP:conf/nips/TancikSMFRSRBN20, DBLP:conf/nips/SitzmannMBLW20, aumentado2022representing, shen2021tetrahedra, yenamandra2024fire, wang2023neuralhessian}. 
To improve the shape modeling capability, researchers inject local-aware designs. For example, DeepLS~\cite{DBLP:conf/eccv/ChabraLISSLN20} integrates voxel grids and local MLPs to decode geometric shapes. Another line of work explores hierarchical representations where the local surfaces are divided unevenly~\cite{10.1145/882262.882293, wang2018adaptive, liu2021octsurf, tang2021octfield, 10.1145/3528223.3530087, takikawa2021nglod}, leveraging Octree. For example, Multilevel Partition of Pnity~\cite{10.1145/882262.882293} (MPU) blends parametric implicit surface patches into a global implicit surface. DOGNet uses dual-octree designs for neural MPU. 
The contribution of \modelname{} is perpendicular to these methods. \modelname{} still works on evenly divided voxels. However, instead of resolving high-frequency details of local shapes by using higher local resolution, \modelname{} proposes the Coordinate Field to reduce the spatial complexity. This is motivated by the analysis result that local geometric shapes are highly compressive under suitable coordinate frames. 


The idea of a coordinate field is related to several existing approaches. For example, MVP~\cite{10.1145/3450626.3459863} introduced oriented boxes for 3D face synthesis. However, in our setting, the variations in geometry and topology are much more significant than those of 3D human faces. 
Another relevant work is LDIF~\cite{DBLP:conf/cvpr/JiangSMHNF20}, which transforms a 3D point in the local coordinate system of each primitive to decode the iso-value of each shape. However, LDIF uses a fixed coordinate frame for each primitive. In contrast, the coordinate field varies spatially in \modelname{} and can be optimized, allowing us to capture detailed variations of the parts flexibly.
Moreover, \modelname{} is based on a rigorous analysis of the expressivity of SDF and SDF learning. A follow-up approach~\cite{DBLP:conf/cvpr/Zheng0DL21} uses a warping field to transform a 3D point into a canonical space of specific categories. In contrast, \modelname{} is category-agnostic, benefiting from the use of local shapes.


On the learning side, many approaches show that MLPs are expressive and that their performance depends on the loss of training. For example, SAL~\cite{DBLP:conf/cvpr/AtzmonL20} and SALD~\cite{DBLP:conf/iclr/AtzmonL21} show the importance of integrating normal losses to capture geometric features. SIREN~\cite{DBLP:conf/nips/SitzmannMBLW20} introduced other regularization losses to improve the quality of implicit representations learned. Although these approaches focus on local shape details, \modelname{} focuses on network design using coordinate frames. The \modelname{} approach is orthogonal to the encoding schemes. 

%\vspace{-0.05in}
\noindent\textbf{Hybrid 3D Representations.} Each 3D representation has fundamental advantages and limitations from the machine learning and representation perspective. For example, implicit representations allow flexible topologies, whereas explicit representations are easier to edit. Therefore, hybrid 3D representations, which aim to add the strength of different 3D representations for representation learning, have received a lot of attention. The main stream in hybrid 3D representations sequentially applies hybrid 3D models~\cite{10.1145/3072959.3073637, 10.1145/3526212, DBLP:conf/iccv/Gkioxari0M19, 10.1145/3355089.3356488, DBLP:conf/nips/ShenGYLF21}. For example, GRASS~\cite{10.1145/3072959.3073637} combines a part-based representation to capture geometric structures of 3D shapes and a volumetric representation per part to capture geometric details of the parts. DSG-Net~\cite{10.1145/3526212} employs part-based deformations to capture geometric details of the part. Other examples~\cite{DBLP:conf/iccv/Gkioxari0M19, 10.1145/3355089.3356488, DBLP:conf/nips/ShenGYLF21, DBLP:conf/eccv/ChabraLISSLN20, hong2023lrm, chan2022efficient, jiang2022few, jiang2023leap} combine explicit graph, mesh, voxel and triplane representations with implicit volumetric representations to encode geometry details. \modelname{} is relevant to this series of approaches, where it combines voxel grids to encode global shapes and an implicit representation to decode local geometric details. The novelty of \modelname{} is that the local module employs a coordinate frame representation and enforces the prior knowledge that the local shape is roughly a low-complexity polynomial surface in the coordinate system defined by normal and principal directions. 


\noindent\textbf{Coordinate Field Optimization.} The task of computing the proposed cell-based coordinate field is related to the problem of vector-field and frame-field design on meshes, where we want to ensure that the coordinate field is smooth and consistent among adjacent cells, and where we want the normal and tangent directions of each coordinate frame to align with the local fitting results if the fitting results are highly confident. This problem was studied in~\cite{10.1145/1183287.1183297}, which introduced a global optimization framework to compute a global vector field on a triangular mesh. Several more recent approaches have developed improved formulations for vector field optimization~\cite{10.1145/2816795.2818078,10.1145/2461912.2462005} and extensions to frame field optimization~\cite{10.1145/2601097.2601179,10.1145/2980179.2982408}. We refer to \cite{10.1145/3084873.3084921} for surveys on this topic. Rather than solving a global optimization problem to compute the coordinate field, the learning of the coordinate field in \modelname{} is driven by learning a compressive MLP. 
%We also introduce a novel loss to address uncertain principal directions and recover suitable orientations of principal directions. 



